Influence of Connectivity on Dung Beetle Communities

### 2025-03-05 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
updates

# total no of spp
total_spp<-btl_data %>% 
  select(-sum) %>% 
  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
  group_by(species) %>%
  summarize(n=sum(n, na.rm=TRUE)) %>%
  arrange(desc(n)) %>%
  mutate(species = reorder(species, desc(n))) %>% 
  summarise(n_spp=n_distinct(species))
```

@@

Study site

@@

Dung beetle sampling

@@

Analyses

@@

  - lets hammer this out 
  - modeling? 
    - glmm with poisson dist reccomended by julian
  - beta, abundance, biomass? per site
  - species list by sampling blocks (anything with this?)
  - habitat preference 
  - rarefaction

@@

most_common<-spp_table_data %>% select(Species) %>% slice(1)
most_common_n<-spp_table_data %>% select(N) %>% slice(1)
```

Overall, `r total_N` total dung beetles were collected belonging to N = `r total_spp` species. The most dominant species were _`r most_common`_ (N = `r most_common_n`), Ateuchus lecontei (n=1115), and Phanaeus igneus (n=958) (Table 1). All but one species was found in matrix plots. Four species (Onthophagus striatulus, Onthophagus concinnus, Geotrupes Blackburnii, Onthophagus truberculifrons) were the only species not representative in every patch type. We had two singleton species, Geotrupes Blackburnii appearing only once in winged, and Onthophagus tuberculiofrons  in matrix.

@@

```{r Figure 1, echo=FALSE,message = FALSE,warning=FALSE}
library(iNEXT)

btl_sums_patch<-btl_data %>%
  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
  group_by(species,patch) %>%
  mutate(patch=case_when(
    patch == "c" ~ "Corridor",
    patch == "m" ~ "Matrix",
    patch == "w" ~ "Winged",
    patch == "r" ~ "Rectangle",
    .default = as.character(patch))) %>%
  summarize(n=sum(n, na.rm=TRUE)) %>%
  arrange(patch,desc(n)) %>% 
  rename(sp_code=species,
         n_patch=n) %>% 
  pivot_wider(
    names_from = patch,
    values_from = c(n_patch))

@@

>>>>

# total no of spp
total_spp<-btl_data %>% 
  select(-sum) %>% 
  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>% 
  group_by(species) %>%
  summarize(n=sum(n, na.rm=TRUE)) %>%
  arrange(desc(n)) %>%
  mutate(species = reorder(species, desc(n))) %>% 
  summarise(n_spp=n_distinct(species))

# 
# most_common<-spp_table_data %>% select(Species) %>% slice(1)
# most_common_n<-spp_table_data %>% select(N) %>% slice(1)

@@

## Study site

@@

## Dung beetle sampling

@@

## Analyses

- species richness

  > glmm with poisson dist reccomended by julian

- species diversity

  > glmm with poisson dist reccomended by julian

- functional diversity

  > 1st pass: roller, tunneler, dweller  
  > after all done: morphological measurements like Alonso et al 2022 (see table from decastro 2023, Giménez Gómez 2025)

- abundance (or biomass)

  > abundance of the most common species?  
  > abundance of species caught in all sites?  
  > changes in rank-abundance.

_Reason for this breakdown: _

- richness and diversity are "community level" response. 
- abundance is a species-level response. 
- changes in abundance or rank-abundance  could have implications for ecosystem services if it is connected to finctional group.

@@

<!--   - lets hammer this out  -->
<!--   - modeling?  -->
<!--     - glmm with poisson dist reccomended by julian -->
<!--   - beta, abundance, biomass? per site -->
<!--   - ~species list by sampling blocks (anything with this?)~ -->
<!--   - habitat preference  -->
<!--   - rarefaction -->

@@

## Overview

- how many beetles total
- most common species overall
- was. the same species the most common one in all habitats?
- were there any that were caught in all habitats? Any restricted to only one?
- any invasive / exotic species?

Overall, `r total_N` total dung beetles were collected belonging to N = `r total_spp` species. The most dominant species were ______  (N = _____), Ateuchus lecontei (n=1115), and Phanaeus igneus (n=958) (Table 1). All but one species was found in matrix plots. Four species (Onthophagus striatulus, Onthophagus concinnus, Geotrupes Blackburnii, Onthophagus truberculifrons) were the only species not representative in every patch type. We had two singleton species, Geotrupes Blackburnii appearing only once in winged, and Onthophagus tuberculiofrons  in matrix.

@@

@@

# REFERENCES

\smallskip  
::: {#refs}
:::

\newpage
# FIGURES

<!-- ```{r  label = pitfalls, echo = TRUE, fig.cap = "Experimental landscapes indicating location of the pitfall traps.", out.width = '50%'} -->
<!-- knitr::include_graphics('./corridor_docs/eric_ms_thesis/pitfalls.png') -->
<!-- ``` -->

![Experimental landscapes indicating location of the pitfall traps.](pitfalls.png)

```{r Figure_hill, echo=FALSE,message = FALSE,warning=FALSE}
library(iNEXT)
#
# btl_data<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_clean","clean_btl_counts.csv"))
# spp_codes<-read_csv(here("corridor_docs","eric_ms_thesis","ms_data","data_raw","species_codes.csv"))

btl_sums_patch<-btl_data %>%
  pivot_longer(pvin:osyl,names_to = "species",values_to = "n") %>%
  group_by(species,patch,block) %>%
  mutate(patch=case_when(
    patch == "c" ~ "Corridor",
    patch == "m" ~ "Matrix",
    patch == "w" ~ "Winged",
    patch == "r" ~ "Rectangle",
    .default = as.character(patch))) %>%
  summarize(n=sum(n, na.rm=TRUE)) %>%
  arrange(patch,desc(n)) %>%
  rename(sp_code=species,
         n_patch=n) %>%
  mutate(pb=paste(patch,block,sep="-")) %>% 
  ungroup() %>% 
  select(-patch,-block) %>% 
  pivot_wider(
    names_from = pb,
    values_from = c(n_patch))

```
\newpage
